---
title: "Final Project Prediction"
author: "Owen Ling"
output: 
    pdf_document
---

* UW ID: 20904591

Important: follow the structure of the document and **summarize important handlings in the summary section**. The structure here is kept at a minimum, feel free if you have more subsections and details to add. Delete this paragraph in the end.

# Summary

## Preprocessing

### Transformation (if any)
* salary: I log-transformed the response variate salary.\
* The following predictor variables were log-transformed: pts, goal, age, takeaway, giveaway, pim, and hits.\
* date: converted to numerical variable using as.numeric(as.date()) function.\
* the following variables were converted into numerical factors: nat, team, pos.\
* The following variables were converted to 2nd-degree polynomials: gp and pm.\

### New Variables
<!-- List all variables/predictors added to dtrain and dtest  -->
* pts_toi: interaction between points and time on ice.\
* goal_toi: intercation between goal and time on ice.\
* toi_gp: interaction between games played and time on ice.\
* pim_hits: interaction between penalty minutes and hits.\
* pm_toi: interaction between plus-minus and time on ice.\
* take_toi: interaction between takeaway and time on ice.\
* give_toi: interaction between giveaway and time on ice.\
* take_pm: interactio between taekaway and plus_minus.\
* toi_inj: interaction between time on ice and games injured.\
* toi_age: interaction between tim eon ice and age.\
* injure_pts: interaction between games injured and points.\

## Model
Summarize your model here: could be the formula for a smoothing method, random forest, etc.\
\
My model is a random forest model with the following formula: "log(salary)~.", which is all the default variables(some transformed as mentioned above in the "transformation section") and the interaction variables in the "new variables" section.
\

## Model Building
Summarize your model building process, such as model selection strategy for a smoothing method, or  parameter combinations searched for a random forest. Report your minimized error measure. If you have attempted multiple methods, report the minimized error measure for each.


<!-- Details -->
<!-- R code starts, please present both code and output. -->
<!-- Provide descriptions and explanations to justify your operations -->
<!-- Be brief and to the point. -->
<!-- Only details lead to your final model are needed. If you tried other appoaches/models, you can mention them, no details needed for abandoned attempts. -->



# 1.Preprocessing

## 1.1 Loading data
```{r,echo=TRUE}
load("final.Rdata")
```

```{r}
library(moments)
#check the skewness of the distribution for each variable to determine whether log transformation is needed
num_vars = c("cap", "gp", "toi", "pts", "goal", "age", 
             "takeaway", "giveaway", "relCorsi", "pim", "hits", "pm")
par(mfrow = c(2, 2))
for (c in num_vars) {
  hist(dtrain[[c]], main = paste("hist of", c), xlab = paste("skewness:", skewness(dtrain[[c]])))
}
```

```{r}
set.seed(444)
library(randomForest)
library(ranger)
library(caret)

  dtrain$date = as.numeric(as.Date(dtrain$date))
  
  #skewness:
  dtrain$pts = log(dtrain$pts+1)
  dtrain$goal = log(dtrain$goal+1)
  dtrain$age = log(dtrain$age)
  dtrain$takeaway = log(dtrain$takeaway+1)
  dtrain$giveaway = log(dtrain$giveaway+1)
  dtrain$pim = log(dtrain$pim+1)
  dtrain$hits = log(dtrain$hits+1)
  
  #categorical->numbers
  dtrain$nat = as.integer(factor(dtrain$nat))
  dtrain$team = as.integer(factor(dtrain$team))
  dtrain$pos = as.integer(factor(dtrain$pos))
  
  #interactions
  dtrain$pts_toi= dtrain$pts*dtrain$toi
  dtrain$goal_toi = dtrain$goal*dtrain$toi
  dtrain$toi_gp = dtrain$toi*dtrain$gp
  dtrain$pim_hits =  dtrain$pim*dtrain$hits
  dtrain$pm_toi = dtrain$pm*dtrain$toi
  
  dtrain$take_toi = dtrain$takeaway*dtrain$toi
  dtrain$give_toi = dtrain$giveaway*dtrain$toi
  dtrain$take_pm = dtrain$takeaway*dtrain$pm
  dtrain$toi_inj= dtrain$toi*dtrain$injure
  dtrain$toi_age = dtrain$toi*dtrain$age
  
  
  dtrain$injure_pts=dtrain$injure*dtrain$pts
```
\
\
```{r}
#plot each variable against salary to examine each individual relationship in detail, 
#in order to determine whether polynomial/exponential transformations are needed.
par(mfrow = c(2, 2))
for (c in num_vars) {
  plot(log(dtrain$salary)~dtrain[[c]], main = paste0("plot of log(salary)~", c)
       , xlab = paste("variable:", c), data = dtrain)
}
```
\
```{r}
#polynomials
  dtrain$gp = poly(dtrain$gp, degree=2)
  dtrain$pm = poly(dtrain$pm, degree= 2)
```
\
# 2.Model Building
```{r}
set.seed(444)
train_control <- trainControl(method = "cv", number = 5, search = "grid")
  fit_test <- train(
    log(salary) ~ .,
    data = dtrain,
    method = "ranger", #this is another function for random forest.
    trControl = train_control,
    tuneLength = 50,
    verbose=1,
    importance = "permutation"
  )

print(fit_test)
fit_test$bestTune
imp = varImp(fit_test)
print(imp)
plot(varImp(fit_test), top = 30)
```
\
I saw that pts, goal, and toi had the most impact on improving the accuracy of the model, so I created many interactions involving these 3 variables that I found to be meaningful.
\
# 3.Evaluation
<!-- Make sure your code runs without error, and record the running time -->
```{r,echo=TRUE}
rm(list=ls())
# first set working directory to the data location
load("final.Rdata")
# final.Rdata contains dtrain and dtest

# load user-defined FinalModel function
source("20904591.R")
# call the function
tmp <- system.time(res <- FinalModel(dtrain, dtest))
# time used, this is the "Running_time" you report in your R source file
tmp[3]

dim(res)
# should be 400 by 2

head(res)
```